---
title: "Exercise_2_Completed"
author: "Daniel T Citron"
date: '2023-07-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up
```{r Load required Libraries}
require("deSolve") # numerical integration package
require("tidyverse") # packages for handling and reshaping data
require("reshape2")
require("magrittr")
require("stats") # statistics package
```

```{r Define SIR model function to be integrated}
sir_model <- function(time, state, parameters) {
  # Unpack Parameters
  with(as.list(c(state,parameters)), {
      # Calculate total population size
      N <-  S + I + R
      
      # Calculate force of infection
      lambda <- beta * I / N
      
      # Calculate derivatives
      dS <- -lambda * S
      dI <- lambda * S - gamma * I
      dR <- gamma * I
      
      # Return derivative
      return(list(c(dS, dI, dR)))
    }
  )
}
```


# Exercise 2: Behavior of the SIR Model

In this exercise we will explore the behavior and properties of the SIR model. For a range of parameter values of beta and gamma, find the size of the epidemic. (This code block may take a minute to run.)
```{r}
# Initial state values
initial_state_values <- c(S = 9999,
                         I = 1,
                         R = 0
)

# Define a list of time steps 
times = seq(from = 0, to = 1000, by = .1)

# Loop over a range of beta and gamma, find the size of the epidemic each time
beta.gamma.outbreak <- data.frame(beta = c(), 
                                  gamma = c(), 
                                  outbreak.size= c())
for (beta.value in seq(0,1,.1)){
  for (gamma.value in seq(.1,1.,.1)){
    parameters <- c(beta = beta.value, # per-contact infection rate
                    gamma = gamma.value # recovery rate
                    )
    # Integrate
    sir.output <- as.data.frame(ode(y = initial_state_values, 
                            times = times, 
                            func = sir_model,
                            parms = parameters))
    # Find the total outbreak size
    total.recovered = sir.output[sir.output$I <1,][1,]$R
    
    beta.gamma.outbreak = rbind(beta.gamma.outbreak,
                                data.frame(beta = beta.value, 
                                           gamma = gamma.value, 
                                           outbreak.size = total.recovered)
                                )

  }
}

#beta.gamma.outbreak

```

Plot the results, the log of the outbreak size as a function of beta and gamma.
```{r}
p <- ggplot(data = beta.gamma.outbreak) + 
  geom_point(mapping = aes(x = gamma, y = beta, color = log(outbreak.size)),
             size = 4) + 
  scale_color_gradient2() + 
  labs("Log(Outbreak Size)") + 
  scale_x_continuous(breaks = c(0,.2,.4,.6,.8,1)) + 
  scale_y_continuous(breaks = c(0,.2,.4,.6,.8,1)) + 
  xlab("gamma") + 
  ylab("Beta") +
  theme_bw()
p
```


## Question: What do you notice about how the outbreak size depends on beta and gamma?

**Pause here**

# Exercise 2 part 2

How does the outbreak size change depending on the basic reproductive number R_0 = beta/gamma?

Fix gamma = .2, vary beta from 0 to 1 such that R_0 varies from 0 to 5.
```{r}
# Define a list of time steps 
times = seq(from = 0, to = 1000, by = .1)

# Loop over a range of beta and gamma, find the size of the epidemic each time
r0.outbreak <- data.frame(R_0 = c(), 
                          outbreak.size= c())

gamma.value = .2
for (beta.value in seq(0,1,.05)){
    parameters <- c(beta = beta.value, # per-contact infection rate
                    gamma = gamma.value # recovery rate
                    )
    # Integrate
    sir.output <- as.data.frame(ode(y = initial_state_values, 
                            times = times, 
                            func = sir_model,
                            parms = parameters))
    # Find the total outbreak size
    total.recovered = sir.output[sir.output$I <1,][1,]$R
    
    r0.outbreak = rbind(r0.outbreak,
                        data.frame(R_0 = beta.value/gamma.value, 
                                   outbreak.size = total.recovered)
                        )

}

#r0.outbreak
```
Plot the size of the outbreak against R_0:
```{r}
p <- ggplot(data = r0.outbreak) + 
  geom_line(mapping = aes(x = R_0, y = outbreak.size), size = 2) + 
  xlab("R_0") + 
  ylab("Outbreak size") +
  theme_bw()

p
```

## Question: How does the outbreak size depend on R_0?
What do you notice about how the model behaves as a function of R_0? Does this result change if we pick a different fixed value for gamma?


# Optional exercise: When does the epidemic die out?

Return to the model in exercise 1.  Plot R-effective (R_eff = R_0 * S/N) and the number of infected individuals together

```{r}

# Initial state values
initial_state_values <- c(S = 9999,
                         I = 1,
                         R = 0
)
# Parameter values
parameters <-  c(beta = 1, # per-contact infection rate
                 gamma = .2 # recovery rate
)
# Time
times = seq(from = 0, to = 50, by = .1)

# Integrate
sir.output <- as.data.frame(ode(y = initial_state_values, 
                            times = times, 
                            func = sir_model,
                            parms = parameters))

# Calculate R_eff = R_0 * S/N
with(as.list(parameters), {
    ggplot(data = sir.output) + 
      geom_line(mapping = aes(x = time, y = beta/gamma * S/(S + I + R)), 
                size = 2 , color = 'black') + 
      geom_line(mapping = aes(x = time, y = I/(S+I+R)), 
                size = 2, color = 'red') + 
      xlab("Time") + 
      ylab("R_eff") + 
      theme_bw()
  }
)
  
```

## Question: At what time does the epidemic peak occur?
```{r}
sir.output[sir.output$I == max(sir.output$I),]$time
```

Let R_eff = R_0 * S/N. This can be interpreted as the number of secondary cases arising from a single infection in a population which has already been partially infected.

## Question: At the time of the epidemic peak, what is the value of R_eff?
```{r}

with(as.list(c(sir.output[which(sir.output$I == max(sir.output$I)) ,], parameters)), {
    beta/gamma * S/(S + I + R)
})

```

## Question: What causes an epidemic to die out?

Given the value of R_eff at the peak of the epidemic, what does that tell us about what causes the number of infected individuals to start decreasing as the epidemic dies out?
