---
title: "Exercise_0_Completed"
author: "Daniel T Citron"
date: '2024-02-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 0: Introduction to Computational Tools - R

With this notebook, we will walk through how to generate the code I created for Exercise 1. This will also form the basis for the other exercises in the workshop. 

This is not the only way to approach and solve the various exercises in this workshop, so I encourage you to also create solutions to the exercises on your own.

## Import Libraries
```{r Load required Libraries, echo=FALSE}
require("tidyverse") # packages for handling and reshaping data
require("reshape2")
require("magrittr")
require("stats") # statistics package
```

The first thing that we will do together in this workshop is to be able to the Ordinary Differential Equation implementation of the Susceptible-Infected-Recovered model. To do this, we are going to use the `ode` function:

```{r Load ode solver, echo = FALSE}
require("deSolve") # numerical integration package
```

What does the function do? Check the documentation
```{r}
?deSolve::ode
```

The important arguments for `ode` are:
  * `y` - initial state values, a list
  * `times` - a list of times where we evaluate the ODE
  * `func` - a function which returns the derivatives in your differential equation
  * `parms` - a list of parameters
To summarize, we pass our differential equation (`func`), a set of initial conditions (`y`), a set of time steps (`times`), and parameters (`parms`) into the function `ode`. The function will numerically integrate the differential equation and return values for that solution at each of the specified time steps.

Let's create an array of initial conditions. You can use the create list ("concatenate") function `c()` to make an array, with elements separated by commas `c(1,2,3)`. The initial conditions are that at time=0 there are 9999 Susceptible individuals, 1 Infected individual, and 0 Recovered individuals:
```{r}
# Uncomment and fill in the initial conditions below
#initial_state_values = c(S_0, I_0, R_0)
```

Let's create an array of time steps where we will evaluate the differential equation. Use the function `seq` to create a sequence of time steps between time=0 and time=50.1, with an interval of 0.1
```{r}
# Define a list of time steps
?seq
# Uncomment and fill in the code below
# times = seq()
```

Let's create an array of arguments `c(Beta, Gamma)` to pass to our differential equation
```{r}
# Uncomment and fill in the values below
#parameters = c(Beta, Gamma)
```

Finally, we need to create a function which evaluates the derivatives of each of our state variables, our differential equation. 

It should take in the following things as arguments

* Array of current state variable values, ordered the same as the initial_state_variables above
* Current time
* Parameter values

It should return an array of derivatives `c(dS, dI, dR)`, ordered the same as the initial_state_variables above

Complete the code below where indicated:
```{r}
sir_model <- function(time, state, parameters) {
  # Unpack Parameters Beta, Gamma

  
  # Unpack current state variables S, I, R

    # Calculate total population size
    # N <- 
    
    # Calculate force of infection
    # lambda <- 
    
    # Calculate derivatives
    # dS <- 
    # dI <- 
    # dR <- 
    
    # Return derivative
    return(list(c(dS, dI, dR)))
}
```


Now, use the `ode` function to integrate:
```{r}
sir.output <- as.data.frame(ode(y = initial_state_values, 
                            times = times, 
                            func = sir_model,
                            parms = parameters))
```

Create a quick plot
```{r}
ggplot(data = sir.output) + 
  geom_line(mapping = aes(x = time, y = S), 
            color = 'blue', size = 2) + 
  geom_line(mapping = aes(x = time, y = R), 
            color = 'black', size = 2) + 
  geom_line(mapping = aes(x = time, y = I), 
            color = 'red', size = 2)

```

