---
title: "Exercise_2_Pseudocode"
author: "Daniel T Citron"
date: '2023-07-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up
```{r Load required Libraries}
require("deSolve") # numerical integration package
require("tidyverse") # packages for handling and reshaping data
require("reshape2")
require("magrittr")
require("stats") # statistics package
```

```{r Define SIR model function to be integrated}
sir_model <- function(time, state, parameters) {
  # Unpack Parameters

  # Calculate force of infection
  #lambda <- 
  
  # Calculate derivatives
  #dS <-
  #dI <- 
  #dR <- 
  
  # Return derivative
  return(list(c(dS, dI, dR)))
}
```


# Exercise 2: Behavior of the SIR Model

In this exercise we will explore the behavior and properties of the SIR model. For a range of parameter values of beta and gamma, find the size of the epidemic. (This code block may take a minute to run.)
```{r}
# Initial state values
initial_state_values <- c(S = 9999,
                         I = 1,
                         R = 0
)

# Define a list of time steps 
#times = 

# Loop over a range of beta and gamma, find the size of the epidemic each time
for (beta.value in seq(0,1,.1)){
  for (gamma.value in seq(.1,1.,.1)){
    # Integrate
    
    # Find the total outbreak size

    # Return total outbreak size vs. beta and gamma
    
  }
}

#beta.gamma.outbreak

```

Plot the results, the log of the outbreak size as a function of beta and gamma.
```{r}

```


## Question: What do you notice about how the outbreak size depends on beta and gamma?

**Pause here**

# Exercise 2 part 2

How does the outbreak size change depending on the basic reproductive number R_0 = beta/gamma?

Fix gamma = .2, vary beta from 0 to 1 such that R_0 varies from 0 to 5.
```{r}
# Define a list of time steps 
times = seq(from = 0, to = 1000, by = .1)

# Loop over a range of beta and gamma, find the size of the epidemic each time
gamma.value = .2

for (beta.value in seq(0,1,.05)){
    # Integrate
    
    # Find and return the total outbreak size
  
}
```
Plot the size of the outbreak against R_0:
```{r}

```

## Question: How does the outbreak size depend on R_0?
What do you notice about how the model behaves as a function of R_0? Does this result change if we pick a different fixed value for gamma?


# Optional exercise: When does the epidemic die out?

Return to the model in exercise 1. Plot R-effective (R_eff = R_0 * S/N) and the number of infected individuals together

```{r}

# Initial state values
initial_state_values <- c(S = 9999,
                         I = 1,
                         R = 0
)
# Parameter values
parameters <-  c(beta = 1, # per-contact infection rate
                 gamma = .2 # recovery rate
)
# Time
times = seq(from = 0, to = 50, by = .1)

# Integrate

# Calculate R_eff = R_0 * S/N
```

## Question: At what time does the epidemic peak occur?
```{r}

```

Let R_eff = R_0 * S/N. This can be interpreted as the number of secondary cases arising from a single infection in a population which has already been partially infected.

## Question: At the time of the epidemic peak, what is the value of R_eff?
```{r}

```

## Question: What causes an epidemic to die out?

Given the value of R_eff at the peak of the epidemic, what does that tell us about what causes the number of infected individuals to start decreasing as the epidemic dies out?
